Viscous Erosion in Porous Media

### 2018-03-26 docs/methods/erosion_methods.tex
NM

With the vanishing rate understood, we now seek to quantify the limiting shape observed in Fig.~\ref{shrink_intface}(a). In particular, consider the corner formed at the front and rear of the body. Locally, we can neglect body curvature and consider the problem of Stokes flow around an infinite wedge; see Fig.~\ref{corner} for a schematic. We seek to predict the wedge's opening angle, $\oangle$, by appealing to the principle that, in long time, erosion uniformly distributes the shear stress along the surface. For a wedge of arbitrary angle, the surrounding flow can be determined exactly in polar coordinates $(r, \theta)$ where $\theta \in [-\thb,\thb]$ and $\thb = \pi - \oangle/2$. (In this section alone, $\theta$ denotes the polar coordinate and not the local tangent angle).

@@

The corresponding stream function is biharmonic, $\Delta^2 \psi = 0$, and the ansatz $\psi = r^{\lambda}f(\theta)$ gives the 4th-order ODE~\cite{poz1997}
\begin{align}
\label{fODE}
f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0.
\end{align}
This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$ and odd symmetry about the horizontal axis, $f(0) = f''(0) = 0$. For arbitrary $\beta$, ODE \eqref{fODE} and the boundary conditions comprise an eigenvalue problem for $\lambda$, whose solution furnishes the stream function and the corresponding shear stress on the wedge surface $\tau = \mu r^{\lambda-2} f''(\thb)$. We, however, seek to determine the particular opening angle that produces {\em uniform} shear stress and therefore set $\lambda = 2$.
The odd-symmetric solution to \eqref{fODE} that corresponds to $\lambda=2$ is given by
\begin{align}
  f(\theta) = B \sin (2 \theta) + C \theta \, .
\end{align}
Imposing the no-slip boundary conditions gives the condition
\begin{align}
  \sin(2 \thb) - 2 \thb \cos(2 \thb) = 0 \, .
\end{align}
This nonlinear equation has solution $\thb \approx 129^\circ$, which implies that the the uniform-stress wedge has an opening angle of $\oangle \approx 102^\circ$.
% Pozrikidis page 239: symmetric flow.

@@

To compare with this analytical prediction, we aim to measure the opening angles formed in our numerical simulations. Section \ref{sec:scaling}, however, shows that the magnitude of the shear stress diverges as the body vanishes, which causes our simulations lose accuracy in the critical moments when the body is expected to approach its limiting shape. To circumvent this problem, we remove the effect of shrinking by forcing $\Vn$ to have zero mean,
\begin{align}
\Vn \leftarrow \Vn - \mean{\Vn} \, .
\end{align}
This modification allows erosion to sculpt the shape as it would normally, but with the area artificially held fixed.

>>>>

With the vanishing rate understood, we now seek to quantify the limiting shape observed in Fig.~\ref{shrink_intface}(a). In particular, consider the corner formed at the front and rear of the body. Locally, we can neglect body curvature and consider the problem of Stokes flow around an infinite wedge; see Fig.~\ref{corner} for a schematic. We seek to predict the wedge's opening angle, $\oangle$, by appealing to the principle that, in long time, erosion uniformly distributes the shear stress along the surface \cite{moo-ris-chi-zha-she2013, moore2017riemann}. For a wedge of arbitrary angle, the surrounding flow can be determined exactly in polar coordinates $(r, \theta)$ where $\theta \in [-\thb,\thb]$ and $\thb = \pi - \oangle/2$ (in this section alone, $\theta$ denotes the polar coordinate and not the local tangent angle).

@@

The corresponding stream function is biharmonic, $\Delta^2 \psi = 0$, and the ansatz $\psi = r^{\lambda}f(\theta)$ gives the 4th-order ODE~\cite{poz1997}
\begin{align}
\label{fODE}
f'''' + 2(\lambda^2 - 2 \lambda + 2)f'' + \lambda^2(\lambda-2)^2 f = 0.
\end{align}
This ODE is subject to no-slip boundary conditions along the wedge $f(\thb) = f'(\thb) = 0$ and odd symmetry about the horizontal axis, $f(0) = f''(0) = 0$. For arbitrary opening angle, ODE \eqref{fODE} and the boundary conditions comprise an eigenvalue problem for $\lambda$, whose solution furnishes the stream function and the corresponding shear stress on the wedge surface $\tau = \mu r^{\lambda-2} f''(\thb)$. We, however, seek to determine the particular opening angle that produces {\em uniform} shear stress and therefore set $\lambda = 2$.
The odd-symmetric solution to \eqref{fODE} that corresponds to $\lambda=2$ is given by
\begin{align}
  f(\theta) = B \sin (2 \theta) + C \theta \, .
\end{align}
Imposing no slip along the wedge surface gives the condition
\begin{align}
  \sin(2 \thb) - 2 \thb \cos(2 \thb) = 0 \, .
\end{align}
This nonlinear equation has solution $\thb \approx 129^\circ$, which implies that the the uniform-stress wedge has an opening angle of $\oangle \approx 102^\circ$.
% Pozrikidis page 239: symmetric flow.

@@

We now aim to measure the opening angle formed in our simulations for comparison with the above prediction. However, Section \ref{sec:scaling} shows that the magnitude of the shear stress diverges as the body vanishes, which causes our simulations lose accuracy in the critical moments when the body is expected to approach its limiting shape. To circumvent this problem, we remove the effect of shrinking by forcing $\Vn$ to have zero mean,
\begin{align}
\Vn \leftarrow \Vn - \mean{\Vn} \, .
\end{align}
This modification allows erosion to sculpt the shape as it would normally, but with the area artificially held fixed.
